Towards a Planckian standard kilogram

### 2018-03-03 ms.tex
formatted in A&A style

\documentclass[prb,preprint]{revtex4-1} 
% The line above defines the type of LaTeX document.
% Note that AJP uses the same style as Phys. Rev. B (prb).

% The % character begins a comment, which continues to the end of the line.

\usepackage{amsmath}  % needed for \tfrac, \bmatrix, etc.
\usepackage{amsfonts} % needed for bold Greek, Fraktur, and blackboard bold
\usepackage{graphicx} % needed for figures
\usepackage{epstopdf}

@@

% Be sure to use the \title, \author, \affiliation, and \abstract macros
% to format your title page.  Don't use lower-level macros to  manually
% adjust the fonts and centering.

\title{A scenario for a Planckian standard kilogram}
% In a long title you can use \\ to force a line break at a certain location.

\author{Michael Ebersold}
\email{michael.ebersold@uzh.ch} 
% If there were a second author at the same address, we would put another 
% \author{} statement here.  Don't combine multiple authors in a single
% \author statement.
\author{Prasenjit Saha}
\email{psaha@physik.uzh.ch}
% Please provide a full mailing address here.
\affiliation{Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland}

% See the REVTeX documentation for more examples of author and affiliation lists.

\date{\today}

\begin{abstract}
As the kilogram assumes its new definition in terms of $h,c,$ and the
standard second, it is interesting to think about a possible future
definition entirely in terms of fundamental constants.  A kilogram
defined as an integer multiple of the Planck mass (or
4594xxxx$\times\sqrt{\hbar c/G}$) would cause no significant change
from its current value.  The practical hurdle to this Planckian
kilogram is measuring $G$ three or four orders of magnitude better
than currently known, so as to find the unknown digits xxxx.  We study
a possible scenario for doing so with an artificial two-body orbit.
The cleanest (admittedly, not the cheapest) option is to place the
system on a Trojan orbit to minimize external gravitational
perturbations, and enclose it in a cavity chilled to liquid-helium
temperatures, to suppress radiation pressure.  Data analysis would
involve marginalizing over at least 15~nuisance parameters (6~for
orbital parameters and 9~for fictitious forces) but that, we show, is
feasible.
\end{abstract}

@@

The reform of SI units to use quantum standards is an ongoing ``news
story'' in physics.\cite{BIPM16}  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For a simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''.\cite{Chao15}
The proposed redefinition of the kilogram is an explicit value for
$h$ and $c$, but we can think of it as
\begin{equation}
\mathrm{kg} \equiv 1.3563925\ldots \times 10^{50} \, \mathrm{s}^{-1}
                   \times \frac{h}{c^2}
\end{equation}
with the exact value to be decided.  As we can see, the kilogram is
coupled to the definition of a second.  That is not a problem in
practice, as the second is standardized orders of magnitude better
than any other unit.  Nonetheless, it is interesting to consider a
logical further step of decoupling the kilogram from other
definitions.  This could be done by introducing the gravitational
constant, since $\sqrt{hc/G}$ has dimensions of mass.
Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum mechanics in a unit system comes from
Planck.\cite{Planck99} In the process of working towards the radiation
law, Planck realized that $c,G,h$ and the Boltzmann constant $k_B$ (in
his notation $G\rightarrow f$, $h\rightarrow b$, $h/k_B\rightarrow a$)
naturally define units for mass, length, time, and temperature.
Indeed, Planck remarked that such units would be meaningful even in
non-human and extraterrestrial cultures.

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ smaller than Planck's definitions.)  Extension to
electrical units is straightforward, but we will not discuss it here.
The Planck length, time, and temperature in themselves are extreme
scales, not probed by any known physical process, and presumably the
regime of quantum gravity.  But extreme scales do not harm if they are
only part of some expression, as they are here.  The Planck mass is
small but microscopically measurable: $\sim20\mu\mathrm{g}$.  This is
of the same order as the uncertainty in the prototype kilogram itself.\cite{BIPM}
Hence the kilogram could be redefined as an integer multiple of the
Planck mass, without significantly perturbing its value.

@@

To fit for $k$ and the 15 accompanying nuisance parameters, we suggest
using a Markov Chain Monte-Carlo (MCMC) method, which are widely used
for similarly large parameter spaces.  MCMC is a Bayesian approach,
which uses the information provided by observed data about a set of
parameters, to update a prior state of knowledge/ignorance about the
set of parameters to become a posterior state of knowledge/ignorance
about a set of parameters. We will not describe in detail how MCMC
works, but refer the reader to an online tutorial from ESO
PyCoffee\cite{esopycoffee} or an introduction paper from van
Ravenzwaaij et al.\cite{vanRavenzwaaij16}

@@

\begin{table}[h]
	\centering
	\caption{This table shows the ranges in which the parameters numerical values in SI units are estimated to make sense and the value which was chosen to generate data.}
	\begin{ruledtabular}
		\begin{tabular}{l c c}
			% The codes above determine the horizontal alignment in each column.
			% Options are l (left), r (right), c (centered), and p (paragraph).
			% The p option allows an entry to be broken into multiple lines, and
			% therefore requires a width specification, in this case 5 centimeters.
			parameter & range & chosen value \\
			\hline	% horizontal line to separate headings from data
			$k$ & $(8.0052$--$8.0112) \cdot 10^{-10}$ & $8.0088 \cdot 10^{-10}$ \\
			$\omega_x$ & $\pm 1.0 \cdot 10^{-8}$ & $0.02 \cdot 10^{-8}$ \\
			$\omega_y$ & $\pm 1.0 \cdot 10^{-8}$ & $-0.03 \cdot 10^{-8}$ \\
			$\omega_z$ & $(1.95$--$2.05) \cdot 10^{-7}$ & $1.996 \cdot 10^{-7}$ \\
			$C_{xx}$ & $0$--$5.0 \cdot 10^{-14}$ & $3.8 \cdot 10^{-14}$ \\
			$C_{yy}$ & $\pm 10^{-14}$ & $0.3 \cdot 10^{-14}$ \\
			$C_{zz}$ & $\pm 10^{-14}$ & $-0.2 \cdot 10^{-14}$ \\
			$C_{xy} = C_{yx}$ & $\pm 10^{-14}$ & $0.3 \cdot 10^{-14}$ \\
			$C_{yz} = C_{zy}$ & $\pm 10^{-14}$ & $0.1 \cdot 10^{-14}$ \\
			$C_{xz} = C_{zx}$ & $\pm 10^{-14}$ & $-0.4 \cdot 10^{-14}$ \\
			$x_0$ & $0.15 \pm 10^{-7}$ & $0.15$ \\
			$y_0$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$z_0$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$v_{x_0}$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$v_{y_0}$ & ($7.30$--$7.31) \cdot 10^{-7}$ & $7.30698296 \cdot 10^{-5}$ \\
			$v_{z_0}$ & $\pm 10^{-7}$ & $0.0$ \\
		\end{tabular}
	\end{ruledtabular}
	\label{tab:parameters}
\end{table}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=1.0\textwidth]{img/4cornerfit.png}
	\caption{Corner plot of the posterior for 4 parameters: $x_0$, $v_{y_0}$, $\omega_z$ and $k$. Indicated with blue lines are the values we calculated the data from. The dashed lines in the histograms along the diagonal represent the median and the 0.16 and 0.84 percentiles respectively.}
	\label{fig:plotCorner4}
\end{figure}

@@

\begin{thebibliography}{99}
\bibliography{ms.bib}
\end{thebibliography}

>>>>

%\documentclass[referee]{aa} % for a referee version
\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style
%\documentclass{aa}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{txfonts}

@@

\title{Towards a Planckian standard kilogram}
\subtitle{}

\author{Michael Ebersold\thanks{michael.ebersold@uzh.ch}%\inst{1}
        \and
        Prasenjit Saha%\inst{2}\fnmsep
       }

\institute{Physik-Institut, University of Zurich, Winterthurerstrasse
  190, 8057 Zurich, Switzerland}

\date{}

\abstract
% context heading (optional)
% {} leave it empty if necessary  
{As the kilogram assumes its new definition in terms of $h,c,$ and the
  standard second, it is interesting to think about a possible future
  definition entirely in terms of fundamental constants.  A kilogram
  defined as an integer multiple of the Planck mass (or
  4594xxxx$\times\sqrt{\hbar c/G}$) would cause no significant change
  from its current value.}
% aims heading (mandatory)
{The practical hurdle to this Planckian kilogram is measuring $G$
  three or four orders of magnitude better than currently known, so as
  to find the unknown digits xxxx.  We study a possible scenario for
  doing so with an artificial two-body orbit.}
% methods heading (mandatory)
{The cleanest (admittedly, not the cheapest) option is to place the
  system on a Trojan orbit to minimize external gravitational
  perturbations, and enclose it in a cavity chilled to liquid-helium
  temperatures, to suppress radiation pressure.}
% results heading (mandatory)
{Data analysis would involve marginalizing over at least 15~nuisance
  parameters (6~for orbital parameters and 9~for fictitious forces)
  but that, we show, is feasible.}
% conclusions heading (optional), leave it empty if necessary 
{}

\keywords{}

@@

The reform of SI units to use quantum standards is an ongoing ``news
story'' in physics.\footnote{International Bureau of Weights and
  Measures, ``DRAFT 9th edition of the SI Brochure", BIPM.}  At the
heart of the reform are two quantum phenomena that appear on
macroscopic scales: the Josephson effect makes $h/e$ measurable as a
voltage-to-frequency ratio, while the quantum Hall effect makes
$h/e^2$ a standard resistance (the von~Klitzing constant).  These
enable electrical units to be defined in terms of $h$ and $e$.  A
further ingenious instrument, known as a watt balance, precisely
compares electrical and mechanical power, thus enabled mass to be
standardized in terms of $h$.  For a simple introduction to the new
SI, we refer the reader to the article on the ``LEGO watt
balance''.\citep{Chao15} The proposed redefinition of the kilogram is
an explicit value for $h$ and $c$, but we can think of it as
\begin{equation}
\mathrm{kg} \equiv 1.3563925\ldots \times 10^{50} \, \mathrm{s}^{-1}
                   \times \frac{h}{c^2}
\end{equation}
with the exact value to be decided.  As we can see, the kilogram is
coupled to the definition of a second.  That is not a problem in
practice, as the second is standardized orders of magnitude better
than any other unit.  Nonetheless, it is interesting to consider a
logical further step of decoupling the kilogram from other
definitions.  This could be done by introducing the gravitational
constant, since $\sqrt{hc/G}$ has dimensions of mass.
Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum mechanics in a unit system comes from
Planck.\cite{Planck99} In the process of working towards the radiation
law, Planck realized that $c,G,h$ and the Boltzmann constant $k_B$ (in
his notation $G\rightarrow f$, $h\rightarrow b$, $h/k_B\rightarrow a$)
naturally define units for mass, length, time, and temperature.
Indeed, Planck remarked that such units would be meaningful even in
non-human and extraterrestrial cultures.

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ smaller than Planck's definitions.)  Extension to
electrical units is straightforward, but we will not discuss it here.
The Planck length, time, and temperature in themselves are extreme
scales, not probed by any known physical process, and presumably the
regime of quantum gravity.  But extreme scales do not harm if they are
only part of some expression, as they are here.  The Planck mass is
small but microscopically measurable: $\sim20\mu\mathrm{g}$.  This is
of the same order as the uncertainty in the prototype kilogram
itself.\footnote{\url{<https://www.bipm.org/utils/common/pdf/cms_m.pdf>}}
Hence the kilogram could be redefined as an integer multiple of the
Planck mass, without significantly perturbing its value.

@@

To fit for $k$ and the 15 accompanying nuisance parameters, we suggest
using a Markov Chain Monte-Carlo (MCMC) method, which are widely used
for similarly large parameter spaces.  MCMC is a Bayesian approach,
which uses the information provided by observed data about a set of
parameters, to update a prior state of knowledge/ignorance about the
set of parameters to become a posterior state of knowledge/ignorance
about a set of parameters. We will not describe in detail how MCMC
works, but refer the reader to an online tutorial from ESO
PyCoffee\footnote{\url{<http://eso-python.github.io/ESOPythonTutorials/ESOPythonDemoDay8_MCMC_with_emcee.html>}}
or an introduction paper from van
Ravenzwaaij et al.\cite{vanRavenzwaaij16}

@@

\begin{table}%[ht]
	\centering
	\caption{This table shows the ranges in which the parameters numerical values in SI units are estimated to make sense and the value which was chosen to generate data.}
%	\begin{ruledtabular}
		\begin{tabular}{l c c}
			% The codes above determine the horizontal alignment in each column.
			% Options are l (left), r (right), c (centered), and p (paragraph).
			% The p option allows an entry to be broken into multiple lines, and
			% therefore requires a width specification, in this case 5 centimeters.
			parameter & range & chosen value \\
			\hline	% horizontal line to separate headings from data
			$k$ & $(8.0052$--$8.0112) \cdot 10^{-10}$ & $8.0088 \cdot 10^{-10}$ \\
			$\omega_x$ & $\pm 1.0 \cdot 10^{-8}$ & $0.02 \cdot 10^{-8}$ \\
			$\omega_y$ & $\pm 1.0 \cdot 10^{-8}$ & $-0.03 \cdot 10^{-8}$ \\
			$\omega_z$ & $(1.95$--$2.05) \cdot 10^{-7}$ & $1.996 \cdot 10^{-7}$ \\
			$C_{xx}$ & $0$--$5.0 \cdot 10^{-14}$ & $3.8 \cdot 10^{-14}$ \\
			$C_{yy}$ & $\pm 10^{-14}$ & $0.3 \cdot 10^{-14}$ \\
			$C_{zz}$ & $\pm 10^{-14}$ & $-0.2 \cdot 10^{-14}$ \\
			$C_{xy} = C_{yx}$ & $\pm 10^{-14}$ & $0.3 \cdot 10^{-14}$ \\
			$C_{yz} = C_{zy}$ & $\pm 10^{-14}$ & $0.1 \cdot 10^{-14}$ \\
			$C_{xz} = C_{zx}$ & $\pm 10^{-14}$ & $-0.4 \cdot 10^{-14}$ \\
			$x_0$ & $0.15 \pm 10^{-7}$ & $0.15$ \\
			$y_0$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$z_0$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$v_{x_0}$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$v_{y_0}$ & ($7.30$--$7.31) \cdot 10^{-7}$ & $7.30698296 \cdot 10^{-5}$ \\
			$v_{z_0}$ & $\pm 10^{-7}$ & $0.0$ \\
		\end{tabular}
%	\end{ruledtabular}
	\label{tab:parameters}
\end{table}

@@

\begin{figure}%[ht]
	\centering
	\includegraphics[width=1.0\textwidth]{img/4cornerfit.png}
	\caption{Corner plot of the posterior for 4 parameters: $x_0$, $v_{y_0}$, $\omega_z$ and $k$. Indicated with blue lines are the values we calculated the data from. The dashed lines in the histograms along the diagonal represent the median and the 0.16 and 0.84 percentiles respectively.}
	\label{fig:plotCorner4}
\end{figure}

@@

\bibliographystyle{aa}
\bibliography{ms.bib}
